Order Preserving Encryption Schemes: \\ Preventing the Leakage of Distribution\\ 6.857 Final Paper \\ Hrishikesh Joshi, Max Kolysh, Mari Miyachi, John Wang

### 2013-05-14 paper.tex
proofing

\section{Introduction}
An order-preserving encryption scheme utilizes an encryption algorithm to produce ciphertexts that preserve the numerical ordering of the plaintexts. This allows an untrusted server to perform order comparison, for the purpose of sorting or range checks, among other applications. The ideal security goal for an order-preserving encryption scheme, IND-OCPA, is to reveal only information regarding the order of the plaintext values. The first ideal-security order-preserving encryption scheme was proposed by Raluca et al (2013) and called a mutable order-preserving encoding, or mOPE.

@@

We present the concept of structural expectation equivalence, which is a mapping from all possible OPE encodings of one scheme to another, with the probability of obtaining each encoding in the first scheme is equal to the probability of obtaining the corresponding encoding in the second scheme. We compare DCS and RTOPE, to prove that the two schemes are structurally expectation equivalent. This equivalence relationship, we show, also proves that both schemes must be IND-CPDA secure.

To understand the performance and equivalence of our two schemes, we implemented both under an honest server model, and evaluated the number of node rewrites using a range of benchmarks.

The rest of the paper is organized as follows. II discusses related work. III presents DCS in more detail and IV describes IND-CPDA. V presents RTOPE in more detail. VI compares the two schemes, DCS and RTOPE, and proves their equivalence. Finally VII concludes.

@@

Research surrounding OPE schemes has made significant progress in recent years. The most recent contribution of the mOPE scheme by Raluca et al (2013) achieves ideal IND-OCPA security, and thus, our schemes are built upon mOPE.

@@

Raluca et al (2013) provides a new security definition, same-time OPE security, which dictates that an adversary can learn the order only for values stored in the application at the same time. This definition is stronger than IND-OCPA. Our security guarantee of IND-CPDA provides that the order only, not the distribution, for ciphertexts is leaked to an adversary.

@@

In order to solve the problem of leaking the distribution of the plaintext values, we have developed the Distribution Confidentiality Scheme (DCS). This scheme uses Raluca et al (2013)'s original mOPE scheme, but adds an additional layer of security. We show that the schemes are comprable in terms of performance.

@@

In the basic version of the scheme, let us assume that we want to encrypt some plaintext value $x \in \mathrm{Z}$ and send it to the server like in Raluca et al (2013)'s mOPE scheme. For DCS, there is an additional parameter $\rho$ which will govern the probability with which the server sees a repeated value. We will generate a random value $r \in [0,1)$ which contains $\rho$ bits, denoted $r_{\rho}$, and append it to the end of $x$. The resulting value $x' = x + r_{\rho}$ will be encrypted with DET. Raluca et al (2013)'s mOPE scheme will work the same as before, but now acting on $x'$.

Inserting in this scheme is simple, since nothing really has to be changed from Raluca et al (2013)'s mOPE scheme. The client will take the value $x$ to insert, create $x'$, then encrypt $x'$ to obtain a ciphertext. Inserting this ciphertext will involve moving down the server's binary search tree. The insertion process is started by the client acquiring the root node. The client then checks to see if $x'$ is larger or smaller than the root's decrypted value. This informs the client as to the direction in which it should move. Once it arrives at a leaf node, the client will insert the new ciphertext in the correct direction.

@@

    We note that there are three cases. First, if $x$ is not repeated (i.e. occurs in the tree only once), then the proposition is trivially true. Second, if $x$ is repeated, and all the repeated values are in the right subtree of the least common ancestor, then one can choose the value of $x$ which has the largest height in the binary search tree. This defines a subtree containing all values of $x$ and which is also rooted at $x$. The same idea follows when all the repeated values of $x$ are in the left subtree. Finally, if the repeated values of $x$ are in both the left and right subtree of the least common ancestor $l$, then we know that $x \leq l$ and $x \geq l$ by the properties of a binary search tree. This implies that $l = x$, whichcompletes the proof.
  \end{proof}

@@

  We presented DCS and RTOPE, two order-preserving encryption schemes that achieve ideal IND-CPDA security, where an adversary learns nothing about the distribution of the plaintext values based on the ciphertexts, and only the order of the elements. We introduce the notion of structurally expectation equivalence, and show that DCS and RTOPE satisfy this equivalence relationship. Finally, we show that DCS and RTOPE achieve comprable performance.

>>>>

\section{Introduction}
An order-preserving encryption scheme utilizes an encryption algorithm to produce ciphertexts that preserve the numerical ordering of the plaintexts. This allows an untrusted server to perform order comparison, for the purpose of sorting or range checks, among other applications. The ideal security goal for an order-preserving encryption scheme, IND-OCPA, is to reveal only information regarding the order of the plaintext values. The first ideal-security order-preserving encryption scheme was proposed by Popa et al (2013) and called a mutable order-preserving encoding, or mOPE.

@@

We present the concept of structural expectation equivalence, which is a mapping from all possible OPE encodings of one scheme to another, with the probability of obtaining each encoding in the first scheme is equal to the probability of obtaining the corresponding encoding in the second scheme. We compare DCS and RTOPE, to prove that the two schemes are not structurally expectation equivalent.

To better understand the performance and equivalence of our two schemes, we implemented both with a client-server framework that supports inseration and query functionality. We evaluated the number of OPE re-encodings that resulted over a range of benchmark inputs. These inputs also had varying amounts of repetition, to better evaluate our schemes under uneven plaintext distributions.

The rest of the paper is organized as follows. II discusses related work. III presents DCS in more detail and IV describes IND-CPDA. V presents RTOPE in more detail. VI compares the two schemes, DCS and RTOPE, and proves their equivalence. VII describes the experimental results of our implementation of the two schemes. Finally VIII concludes.

@@

Research surrounding OPE schemes has made significant progress in recent years. The most recent contribution of the mOPE scheme by Popa et al (2013) achieves ideal IND-OCPA security, and thus, our schemes are built upon mOPE.

@@

Popa et al (2013) provides a new security definition, same-time OPE security, which dictates that an adversary can learn the order only for values stored in the application at the same time. This definition is stronger than IND-OCPA. Our security guarantee of IND-CPDA provides that the order only, not the distribution, for ciphertexts is leaked to an adversary.

@@

In order to solve the problem of leaking the distribution of the plaintext values, we have developed the Distribution Confidentiality Scheme (DCS). This scheme uses Popa et al (2013)'s original mOPE scheme, but adds an additional layer of security. We show that the schemes are comprable in terms of performance.

@@

In the basic version of the scheme, let us assume that we want to encrypt some plaintext value $x \in \mathrm{Z}$ and send it to the server like in Popa et al (2013)'s mOPE scheme. For DCS, there is an additional parameter $\rho$ which will govern the probability with which the server sees a repeated value. We will generate a random value $r \in [0,1)$ which contains $\rho$ bits, denoted $r_{\rho}$, and append it to the end of $x$. The resulting value $x' = x + r_{\rho}$ will be encrypted with DET. Popa et al (2013)'s mOPE scheme will work the same as before, but now acting on $x'$.

Inserting in this scheme is simple, since nothing really has to be changed from Popa et al (2013)'s mOPE scheme. The client will take the value $x$ to insert, create $x'$, then encrypt $x'$ to obtain a ciphertext. Inserting this ciphertext will involve moving down the server's binary search tree. The insertion process is started by the client acquiring the root node. The client then checks to see if $x'$ is larger or smaller than the root's decrypted value. This informs the client as to the direction in which it should move. Once it arrives at a leaf node, the client will insert the new ciphertext in the correct direction.

@@

    We note that there are three cases. First, if $x$ is not repeated (i.e. occurs in the tree only once), then the proposition is trivially true. Second, if $x$ is repeated, and all the repeated values are in the right subtree of the least common ancestor, then one can choose the value of $x$ which has the largest height in the binary search tree. This defines a subtree containing all values of $x$ and which is also rooted at $x$. The same idea follows when all the repeated values of $x$ are in the left subtree. Finally, if the repeated values of $x$ are in both the left and right subtree of the least common ancestor $l$, then we know that $x \leq l$ and $x \geq l$ by the properties of a binary search tree. This implies that $l = x$, which completes the proof.
  \end{proof}

@@

  We presented DCS and RTOPE, two order-preserving encryption schemes that achieve ideal IND-CPDA security, where an adversary learns nothing about the distribution of the plaintext values based on the ciphertexts, and only the order of the elements. DCS generates random bits to append to any plaintext value that is encrypted, which requires a modified querying scheme involving localized range queries. RTOPE handles repeated values differently, by appending repeated plaintexts as left or right children on the server's search tree with uniform probability. Again, we created a modified querying algorithm for RTOPE that requires recursive queries once one occurence of the plaintext query is found.

  We proved that DCS and RTOPE are IND-CPDA secure and introduced the notion of structurally expectation equivalence. We also proved the important theorem that if two OPE schemes are structurally expectation equivalent, then one is IND-CPDA secure if and only if the other is IND-CPDA secure as well. We then showed that though DCS and RTOPE do satisfy this structurally expectation equivalence relationship, they are still IND-CPDA secure.

  We noted that though the schemes are not structurally expectation equivalent, the number of OPE re-encodings will converege to the same number, given identical inputs. Finally, this notion was further supported by our experimental results. We tested our implementation of DCS and RTOPE on a range of inputs, and found that the number of OPE re-encodings were of the same order.

  \begin{thebibliography}{9}

\bibitem{adapopa13}
  R.A. Popa, F. Li, N. Zeldovich.
  An Ideal-Security Protocol for Order Preserving Encoding.
  2013.

  \bibitem{Shi07}
  E. Shi, J. Bethencourt, T.-H. H. Chan, D. Song, and A. Perrig.
Multi-dimension range query over encrypted data. In \emph{IEEE
Symposium on Security and Privacy}, 2007.

\end{thebibliography}
